Upper Limits on CN from Falling Evaporating Bodies\\seen towards $\beta$~Pictoris

### 2025-02-06 src/tex/ms.tex
finishing

@@

@@

% text highlighting
\usepackage{soul}
\sethlcolor{yellow}

@@

   \author{M.A. Kenworthy
          \inst{1}
          \and
          E. de Mooij\inst{2}
          \and
          C. Opitom\inst{3}
          \and
          A.\ Brandeker\inst{4}
          \and 
          F. Kiefer\inst{5}
          \and
          A. Fitzsimmons \inst{1}
          }

@@

  \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
   {The young (23 Myr) nearby (19.4 pc) star $\beta$~Pictoris hosts an edge-on debris disk with two gas giant exoplanets in orbit around it.
   %
   Many transient absorption features have been detected in the rotationally broadened stellar lines of the star's chromosphere, and these are thought to be the coma of infalling exocomets crossing the line of sight towards Earth.
}
  % aims heading (mandatory)
   {In the Solar system, the molecule cynaogen (\ce{CN}) and its associated ionic species are one of the most detectable molecules in the coma and tails of comets.
   %
   We perform a search for cyanogen in the spectra of \bp{} to detect or put an upper limit on this molecule's presence in a young, highly active planetary system.}
  % methods heading (mandatory)
   {We divide twenty year's worth of \ac{harps} spectra into those with strong exocomet absorption features, and those with only stellar lines.
   %
   The high signal to noise stellar spectrum normalises out the stellar lines in the exocomet spectra, which are then shifted and stacked on the deepest exocomet absorption features to produce a high signal to noise exocomet spectrum, and search for the \ce{CN} band head using a model temperature dependent cross-correlation template.
   }
  % results heading (mandatory)
   {We do not detect \ce{CN} in our data, and place a temperature and broadening dependent 5$\sigma$ upper limit between 10$^{12}$\,cm$^{-2}$ and 10$^{13}$\,cm$^{-2}$, which is 100 times higher than seen for comets in our Solar System.%No \ce{CN} is seen to a level of XXX ppt, indicating that there is less \ce{CN} in these \ac{feb} than Solar system comets.
   %
   }
  % conclusions heading (optional), leave it empty if necessary 
   %{\ac{feb} are not Solar System comets in composition.}

@@

To the contrary of what is observed in the case of \ac{feb}, most molecular and atomic signatures detected in the coma of Solar system comets are emission lines/bands instead of absorptions.
%
Spectroscopic observations of Solar system comets when they are very close to the Sun are rare due to observational difficulties.
%
The only comet spectroscopically observed closer than 0.2~au from the Sun is C/1965~S1 Ikeya-Seki.
%
%For this comet, it was noticed that the main emission bands/lines detected in its optical spectrum dramatically changed while the comet approached the Sun.
%
%At distances greater than 0.4~au, \ce{CH}, \ce{CN}, \ce{C2}, and \ce{C3} emission bands dominated the spectrum, but at 0.15 au from the Sun, the only molecular emission detected was \ce{CN}, while numerous atomic lines were visible \citep[\ce{[\ion{O}{i}], \ion{Na}{i}, \ion{K}{i}, \ion{Ca}{ii}, \ion{Cr}{i}, \ion{Mn}{i}, \ion{Fe}{i}, \ion{Ni}{i}, \ion{Cu}{i}, \ion{V}{i}; ][]{Dufay65,Thackeray1966,Preston1967,Slaughter1969}.
%%\ion{Ca}{ii} and \ion{Fe}{i} have only been spectroscopically observed via emission lines in the sun-grazing comet C/1965~S1 Ikeya-Seki.
%
For this comet, it was noticed that the main emission bands/lines detected in its optical spectrum dramatically changed while the comets approached the Sun.
%
At distances greater than 0.4~au, \ce{CH}, \ce{CN}, \ce{C2}, and \ce{C3} emission bands dominated the spectrum.
%
However, at 0.15~au from the Sun, the only molecular emission detected was \ce{CN}, while numerous atomic lines were visible \citep[\ion{[O]}{i}, \ion{Na}{i}, \ion{K}{i}, \ion{Ca}{ii}, \ion{Cr}{i}, \ion{Mn}{i}, \ion{Fe}{i}, \ion{Ni}{i}, \ion{Cu}{i}, \ion{V}{i}; ][]{Dufay65,Thackeray1966,Preston1967,Slaughter1969}.
%
Additionally, a tail of \ion{Fe}{i} atoms was proposed to explain a linear feature imaged in the tail of sungrazing comet C/2016~P1 McNaught 2006P1 \citep{Fulle2007} and Na and K were detected for comet C/2011 L4 \citep{Fulle13} at 0.46 au.

On the other hand, \ion{Na}{i} D-line emission has been observed in many comets within $r \sim 1$ au of the Sun, where either the emission was much brighter than the foreground sky emission, or the comet was observed at high spectral resolution enabling separation from the telluric lines \citep{Cremonese02,Schmidt16}.
%
The high efficiency of the \ion{Na}{i} D transition makes it easily detectable, even for relatively low column density.
%
A notable discovery was a spectacular tail of \ion{Na}{i} associated with comet C/1995 O1 Hale-Bopp at $r\sim 1$ au, formed by radiation pressure on the neutral atoms before ionisation \citep{Cremonese1997}. 
Recently, \cite{Hui2023} also demonstarted thet the activity of near-Earth asteroid Phaeton, with a perihelion of only 0.14 au, is associated with emission of sodium.

@@

Of all cometary gas emission, the highest signal to noise normally occurs with the \ce{CN} (0-0) band at $\lambda\simeq 3880$\AA.
%
It is usually the first emission detected in the optical spectrum of a comet as it approaches the Sun, and as been detected in the coma of comets observed at relatively large distances from the Sun \citep[e.g. ][]{Fitzsimmons1996}.
%
The relative abundance in the coma of most comets of \ce{OH/CN}$\simeq500$ \citep{AHearn1995}, together with its high fluorescence scattering efficiency factor of $\sim$ 2.6 photons/sec/molecule at 1~au from the Sun and $\dot{r}\simeq 0$ \citep{Schleicher2010}, gives rise to a clear signature of a cometary gas comae.
%
Historically, the \ce{CN} was believed to be created from photodissociation of \ce{HCN}, which is present in abundances of the order of 1\% relative to water in cometary ices.
%
High-spatial resolution sub-mm mapping of \ce{HCN} distributions supports a nuclear source in some comets \citep{Cordiner2014}.
%
However, it is now known that in many comets a substantial fraction of \ce{CN} is released from sub-micron dust grains in the inner coma of comets \citep[e.g. ][]{Fray2005}.

The \ce{C2} (0-0) band at $\lambda\simeq 5160$\AA \ may be of comparable flux, but has a lower contrast with the underlying dust grain continuum due to its wider bandwidth.
%
The \ce{OH} (0-0) band at $\lambda\simeq 3080$ \AA \ generally has higher flux, but unless observed from space the high terrestrial atmospheric absorption severely diminishes the observed flux.
%
Strong \ce{CO+} emission bands have also been observed in a number of comets \citep[the strongest ones are the (3-0) and (2-0) $\mathrm{A^2\Pi-X^2\Sigma}$; ][]{Biver2018}.
%
However this is normally significantly weaker than the observed \ce{CN} emission, with only less than 20 comets.

Given the ubiquitous signature of \ce{CN} emission in Solar system comets, we have embarked on an archival search for this species in the \ac{feb} around \bp{}.
%
Detection of this species would provide a solid link between cometary bodies in our Solar system and the \bp{} system.

@@

For the study in this paper, we use all the publicly available data from the \ac{harps} and the ESO 3.6m telescope in Chile.
%
%In addition, we complement this with observations with the \ac{uves} at the Very Large Telescope in Chile.

@@

We collated all publicly available HARPS observations of \bp{} obtained between October 2003 and March 2020.
%
Although these observations cover a large baseline, the time between visits varies between 1 day and $\sim$2.2 years.
%
Similarly, the number of observations per epoch spans a wide range, between 1 and 356 exposures, while the exposure times vary between 12 and 900 seconds. This results in an average SNR across order 0 - 10 (order numbers for HARPS will follow the labeling in the fits header -- corresponding to 3780\,$\AA$ and 4075\,$\AA$) per exposure between 1 and 430, and a nightly SNR between 63 and 2000.
%
To avoid potential edge effects from order merging,  we used the E2DS files downloaded directly from the ESO archive for our analysis.

% \hl{Q: Do we want to put an observing log in the appendix?}
% \hl{Q: Do we want to attempt to use the S1D, as they may have less issues with wiggles if the blaze profile was removed.}

%\subsection{UVES observations}

%To supplement the HARPS observations, we use 160 epochs of \ac{uves} observations obtained between 01 April, 2017 and April 18, 2018 under programme 599.C-0751.
%
%The details for these observations are given in \cite{vansluijs2019}, and summarised here.
%
%These observations were all obtained using the same instrument setup.
%
%We used the standard 437+760 mode of \ac{uves}, which uses the \#2 dichroic to allow simultaneous observations in both the red and blue arm.
%
%In this mode, the CD\#2 and CD\#4 cross-dispersers are used in the blue and red-arm, respectively, providing a simultaneous wavelength coverage of $\sim 3730$ \AA{} to $\sim 5000$ \AA{} in the blue arm, and $\sim 5650$ \AA{} to $\sim 9560$ \AA{} in the red arm.
%
%For the remainder of the paper, we will only focus on the data from the blue arm of \ac{uves}, as this covers the strong \ce{CN} (0-0) band-head seen in Solar system comets.
%
%For the blue arm, a 0.3''-wide slit was used to obtain the highest possible resolving power ($\sim$90,000 before an intervention on the instrument in October 2017, and $\sim$100,000 after the intervention).

%The data were reduced using the ESO \ac{uves} pipeline version \hl{6.1.6} through ESO Reflex tool in order to provide as homogeneous a dataset as possible. The flat-field was applied on a pixel-by-pixel basis. \hl{The pipeline was set to output the intermediate products and the blaze-corrected spectra before merging were used for our analysis, as the order-merging could introduce systematics in the overlap region. Optimal extraction was performed, using an empirical profile}

\subsection{Initial data processing}
The analysis was started from the 2d extracted spectra (E2DS files). %, the data from both instruments was processed in similar ways.
%
Each order was treated separately, with the main focus on orders 3 and 7 % for HARPS and \hl{4, and 7} for UVES 
as these orders contain the main \ce{CN} bandhead as well as the \ion{Ca}{ii} H line, which is used to identify the exocomets.
%
All the spectra were transformed into the stellar rest-frame and onto a common wavelength grid using linear interpolation. %For UVES, the \ce{Fe} \hl{NNNN angstrom} line was used to track global wavelength drifts between different exposures which were corrected for at the same time. To measure these, a Gaussian was fit to the line, and the velocity offset from the rest frame position determined.

@@

\subsection{Combining the spectra and flattening them}\label{sect:comb}
After the blaze-variation corrections, all the spectra within a night are combined in order to further boost the signal-to-noise and improve the detection limits. Each individual spectrum was weighted using the average signal-to-noise ratio provided in the header for orders 0 and 11 (3780\,$\AA - $4075$\AA$).%\hl{, and OO and PP} for HARPS and UVES respectively.

%For the spectra within a given night:
%
%\begin{itemize}
%\item We median combine all the nights together.%
%
%\item Bin the spectrum and clip it, interpolate using a cubic spline.
%
%\item Divide this interpolated cubic spline into each of the nightly spectra.
%
%\end{itemize}

@@

After this post-processing the overall (residual) blaze profile and stellar spectrum still need to be removed. Care must be taken to prevent the removal of (potential) narrow features from the circumstellar disk (as seen in e.g., the \ce{Fe} and \ion{Ca}{II} lines). This is done by combining the nightly spectra without significant cometary activity (see Sect.~\ref{sect:FEBid}) and binning this stacked spectrum in wavelength in bins of 21 pixels. Before binning, the spectrum in each bin was MAD (median absolute deviation) clipped to within $\pm$4.5 MAD of the median of the flux in the bin. The clipped average of the fluxes within each bin was used. Subsequently, the overall spectrum of the star and (residual) blaze was determined by cubic interpolation across the binned points, and divided out of the spectrum for each night. As before, this was done separately pre- and post-fiber change.

%Subsequently, we applied a correction for variations in the instrumental blaze by selecting a reference spectrum and dividing all the spectra by this reference. The ratio was then binned, removing outliers, and a polynomial was fit to the data.
%
%For \ac{uves} we used a Nth degree polynomial, while for \ac{harps} we used a 2nd degree polynomial.
%
%This polynomial was then divided from the spectrum under consideration. We note that the blaze correction will also remove non-instrumental effects, including variations in atmospheric transmission and seeing.

%The blaze-corrected spectra taken within a night were averaged together [Should we use a weighted average?].
%
%We note that for \ac{harps}, the 2d-unmerged spectra still contain the overal instrumental blaze-profile, which is (partially) corrected for by using a dedicated blaze-frame obtained from the archive, which removes most, but not all, of this profile.

@@

After the initial data analysis, we proceed to with our search for lines from \ce{CN}.
%
Although we are interested in lines from the exocomets, we first co-add the data from all the individual epochs, in the stellar rest-frame, to obtain a deep search for circumstellar \ce{CN} absorption.
%
These coadded spectra %for \ac{harps} and \ac{uves} 
are shown in Figure~\ref{fig:spec_CN_stellar_frame}.
%
No obvious feature from the \ce{CN} band jumps out, and we need to combine the signal from the lines within the \ce{CN} band to improve the detection limits.

Since the exocomets have a clear radial velocity with respect to \bp{}, and since each exocomet will have its own distinct velocity, we selected the 36 epochs with the strongest exocometary features, a seen in the \ion{Ca}{ii} H line.

\subsection{Identifying exocomets from \ion{Ca}{ii} H line}\label{sect:FEBid}
The comet features in the \ion{Ca}{ii} H-line ($\lambda=3968. \AA$) were identified after removing the overall stellar line-profile and avoiding the absorption from the circumstellar disk by excluding the central 8 km\,s$^{-1}$.
%
The deepest absorption feature in each observing night was selected, and the location and depth stored.
%
For the comet sample we selected features deeper than 60\% %\hl{and 40\%} for HARPS and UVES, respectively, 
while spectra with features below 15\% %for both HARPS and UVES 
were flagged as `low activity'.
%
%An example of the \ion{Ca}{ii} H line for a night with a strong comet and for a night with a weak comet are shown in Figure \hl{XXXX}.
%
The distribution of feature locations and depths is shown in Figure~\ref{fig:FEB_velocity}.

The spectra for the strong comet sample were coadded, and the combined spectrum for the HARPS data is shown in Figure~\ref{fig:spec_CN_comet_frame}.
%
As with the circumstellar \ce{CN}, there are no obvious lines from \ce{CN} visible, and we will need to combine multiple lines to improve the signal-to-noise ratio, and see if there is any \ce{CN} present. 
%
%The spectra with only weak exocometary features were co-averaged in the stellar rest frame in order to create a stellar master spectrum. This master-spectrum was divided out of all the remaining spectra in order to create relative spectra. 
%
%The relative velocities of the exocomets were determined by ......., and the spectra were shifted to the rest-frames of these comments, before being combined.
%
%The combined exocomet aligned spectra for both the \ce{Ca} H\&K lines and \ce{CN} band are shown in Fig.~\ref{fig:coadd_exocomets}. As can be seen no signal is visibly present.
%

\begin{figure}
    \begin{centering}
        \includegraphics[width=\columnwidth]{figures/spec_stack_CN_stellar.pdf}
        \caption{Plot of the stacked spectrum in the stellar restframe (blue) with a CN model for $T=$300K and N=10$^{14}$\,cm$^{-2}$ overplotted in orange. The large signal around 3860\,$\AA$ is due to the residual from a circumstellar \ion{Fe}{I} line.}
        \label{fig:spec_CN_stellar_frame}
        \script{plot_spec_mean.py}
    \end{centering}
\end{figure}

\begin{figure}
    \begin{centering}
        \includegraphics[width=\columnwidth]{figures/spec_stack_CN_comet.pdf}
        \caption{Plot of the stacked spectrum in the stellar restframe (blue) with a CN model for $T=$300K and N=10$^{14}$\,cm$^{-2}$ overplotted in orange. The large signal around 3860\,$\AA$ is due to the residual from a circumstellar \ion{Fe}{I} line.}
        \label{fig:spec_CN_comet_frame}
        \script{plot_spec_mean.py}
    \end{centering}
\end{figure}

@@

%\subsection{Summary of \ce{CN} in comets}

\ce{CN} is one of the brightest optically active species in cometary comae.
%
It is a minor component of the gas coma, with a production rate by number $\sim 1/500$th that of water.

Most studies of Solar system comets simply observe the \ce{CN} molecules via optical fluorescence.
%
From measurements and simple physical arguments the \ce{CN} absorption/emission is optically thin everywhere.
%
This means a flux from part of the coma can immediately be converted to a column density using a photon scattering efficiency for the molecular band.

\ce{CN} in solar system comets comes from two sources.
%
Sublimation of nucleus near-surface ices releases \ce{HCN}, which then photo-dissociates into \ce{CN}.
%
This in turn eventually photo-dissociates into the component atoms.

@@

Since Halley in 1986, it has also been known that dust grains ejected from the nucleus also directly release \ce{CN} into the coma.
%
However, \ce{CN} production in comets is usually dominated by nuclear ice sublimation.

The resulting \ce{CN} coma will have a true space density and apparent column density distribution centrally peaked at the comet nucleus.
%
For Solar system comets, column densities are often converted to an overall production rate $Q_{\ce{CN}}$ molecules/sec via an analytical model known as the ``Haser'' model.
%
For this model there are only 3 parameters; the effective destruction scale length $l_p$ of the parents of \ce{CN} (\ce{HCN}+dust), the photodissociation scale length $l_{\ce{CN}}$ of \ce{CN}, and the gas outflow velocity $v$.
%
The Haser model is somewhat simplistic but it represents the data well for solar system comets observed at typical distances from the SUn (1-3 au).

@@

We take the observed activity of 1P/Halley in 1985/86 as an example for our estimates.
%
Although most comets we observe are far less active than Halley, there are quite a few comets with activity similar to it.
%
Given that we will be dominated by the most active comets at \bp{}, this seems reasonable as something that can be later scaled to the brightest comets such as Hale-Bopp.
%
When Halley was at $\sim 1$ au from the Sun, it was measured to have $Q_{\ce{CN}}\simeq 10^{27}$ molecules/sec using Haser models.
%
The assumed outflow velocity in the models was generally  $0.85 - 1$ \kms{} at this distance.

For solar system comets, the normal Haser scale lengths used \citep{Cochran86,AHearn1995} are:

$$lp=1.7\times10^4r_h^2\:{\rm km}$$

$$l_{\ce{CN}}=2-3\times10^5r_h^2\:{\rm km}$$

@@

The lifetime of \ce{HCN} for photodissociation into \ce{H}+\ce{CN} at 1~au has been calculated to be $(3.2-7.7)\times10^4$ sec, with lifetime of \ce{CN} dissociating into \ce{C}+\ce{N} being $(1.3-3.1)\times10^5$ sec \citep{Huebner92}, depending on solar activity levels.
%
These are consistent with the measured scale lengths above if the parents of \ce{CN} are ejected at $\sim 0.5$ km/sec, and we take into account the excess photodissociation energy giving \ce{CN} an extra kick of 1 \kms{}.

@@

Photodissociation is  driven by the UV photon flux at the comet, which in turn is dominated by Ly-$\alpha$ line flux. From Ly-$\alpha$ measurements in \citet{Landsman93} the intrinsic Ly-$\alpha$ luminosity of A7V stars is a factor of $\sim10$ higher than for G2V stars.
%
Hence the photodissociation lifetime of \ce{CN} will be 10 times smaller i.e. $\sim 2\times10^4$ sec at 1 au.

In steady-state sublimation, the sublimation rate is proportional to the photon flux at the comet. For \bp{} $L=8.7L_\odot$, hence assuming dust desorption also scales this way, Halley would have a larger \ce{CN} production rate of $Q_{\ce{CN}}\sim 8\times10^{27}$ molecules/sec.
%
The total number of \ce{CN} molecules in the coma at 1~au would therefore be

$$N_{\ce{CN}}=Q_{\ce{CN}} \tau_{\ce{CN}}$$

$$N_{\ce{CN}}\sim 8\times10^{27} \cdot  2\times10^4 \sim 2\times10^{32}\:{\rm molecules}$$

%\subsubsection{\ce{CN} Cross-section}

>>>>

\newcommand{\angstrom}{\textup{\AA}}

@@

%\newcommand\pdfmath[1]{\texorpdfstring{$#1$}{#1}}

@@

@@

   \author{M.A. Kenworthy
          \inst{1}
          \and
          E. de Mooij\inst{2}
          \and
          C. Opitom\inst{3}
          \and
          A.\ Brandeker\inst{4}
          \and 
          F. Kiefer\inst{5}
          \and
          A. Fitzsimmons \inst{2}
          }

@@

  \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
   {The young (23 Myr) nearby (19.4 pc) star $\beta$~Pictoris hosts an edge-on debris disk with two gas giant exoplanets in orbit around it.
   %
   Many transient absorption features have been detected in the rotationally broadened stellar lines of the star's chromosphere, and these are thought to be the coma of infalling exocomets crossing the line of sight towards Earth.
}
  % aims heading (mandatory)
   {In the Solar system, the molecule cynaogen (\ce{CN}) and its associated ionic species are one of the most detectable molecules in the coma and tails of comets.
   %
   We perform a search for cyanogen in the spectra of \bp{} to detect or put an upper limit on this molecule's presence in a young, highly active planetary system.}
  % methods heading (mandatory)
   {We divide twenty year's worth of \ac{harps} spectra into those with strong exocomet absorption features, and those with only stellar lines.
   %
   The high signal to noise stellar spectrum normalises out the stellar lines in the exocomet spectra, which are then shifted and stacked on the deepest exocomet absorption features to produce a high signal to noise exocomet spectrum, and search for the \ce{CN} band head using a model temperature dependent cross-correlation template.
   }
  % results heading (mandatory)
   {We do not detect \ce{CN} in our data, and place a temperature and broadening dependent 5$\sigma$ upper limit between 10$^{12}$\,cm$^{-2}$ and 10$^{13}$\,cm$^{-2}$, which is 100 times higher than seen for comets in our Solar System.
   }
  % conclusions heading (optional), leave it empty if necessary 
   %{\ac{feb} are not Solar System comets in composition.}

@@

To the contrary of what is observed in the case of \ac{feb}, most molecular and atomic signatures detected in the coma of Solar system comets are emission lines/bands instead of absorptions.
%
Spectroscopic observations of Solar system comets when they are very close to the Sun are rare due to observational difficulties.
%
The only comet spectroscopically observed closer than 0.2~au from the Sun is C/1965~S1 Ikeya-Seki.
%
For this comet, it was noticed that the main emission bands/lines detected in its optical spectrum dramatically changed while the comets approached the Sun.
%
At distances greater than 0.4~au, \ce{CH}, \ce{CN}, \ce{C2}, and \ce{C3} emission bands dominated the spectrum.
%
However, at 0.15~au from the Sun, the only molecular emission detected was \ce{CN}, while numerous atomic lines were visible \citep[\ion{[O]}{i}, \ion{Na}{i}, \ion{K}{i}, \ion{Ca}{ii}, \ion{Cr}{i}, \ion{Mn}{i}, \ion{Fe}{i}, \ion{Ni}{i}, \ion{Cu}{i}, \ion{V}{i}; ][]{Dufay65,Thackeray1966,Preston1967,Slaughter1969}.
%
Additionally, a tail of \ion{Fe}{i} atoms was proposed to explain a linear feature imaged in the tail of sungrazing comet C/2016~P1 McNaught 2006P1 \citep{Fulle2007} and Na and K were detected for comet C/2011 L4 \citep{Fulle13} at 0.46 au.

On the other hand, \ion{Na}{i} D-line emission has been observed in many comets within $r \sim 1$ au of the Sun, where either the emission was much brighter than the foreground sky emission, or the comet was observed at high spectral resolution enabling separation from the telluric lines \citep{Cremonese02,Schmidt16}.
%
The high efficiency of the \ion{Na}{i} D transition makes it easily detectable, even for relatively low column density.
%
A notable discovery was a spectacular tail of \ion{Na}{i} associated with comet C/1995 O1 Hale-Bopp at $r\sim 1$ au, formed by radiation pressure on the neutral atoms before ionisation \citep{Cremonese1997}. 
Recently, \cite{Hui2023} also demonstrated thet the activity of near-Earth asteroid Phaeton, with a perihelion of only 0.14 au, is associated with emission of sodium.

@@

Of all cometary gas emission, the highest signal to noise normally occurs with the \ce{CN} (0-0) band at $\lambda\simeq 3880$\AA.
%
It is usually the first emission detected in the optical spectrum of a comet as it approaches the Sun, and as been detected in the comae of comets observed at relatively large distances from the Sun \citep[e.g. ][]{Bus1991,Fitzsimmons1996}.
%
The relative abundance in the coma of most comets of \ce{OH/CN}$\simeq500$ \citep{AHearn1995}, together with its high fluorescence scattering efficiency factor of $\sim$ 2.6 photons/sec/molecule at 1~au from the Sun and $\dot{r}\simeq 0$ \citep{Schleicher2010}, gives rise to a clear signature of a cometary gas coma.
%
Historically, the \ce{CN} was believed to be created from photodissociation of \ce{HCN}, which is present in abundances of the order of 1\% relative to water in cometary ices.
%
High-spatial resolution sub-mm mapping of \ce{HCN} distributions supports a nuclear source in some comets \citep{Cordiner2014}.
%
However, it is now known that in many comets a substantial fraction of \ce{CN} is released from sub-micron dust grains in the inner coma of comets \citep[e.g. ][]{Fray2005}.

The \ce{C2} (0-0) band at $\lambda\simeq 5160$\AA \ may be of comparable flux, but has a lower contrast with the underlying dust grain continuum due to its wider bandwidth.
%
The \ce{OH} (0-0) band at $\lambda\simeq 3080$ \AA \ generally has higher flux, but unless observed from space the high terrestrial atmospheric absorption severely diminishes the observed flux.
%
Strong \ce{CO+} emission bands have also been observed in a number of comets \citep[the strongest ones are the (3-0) and (2-0) $\mathrm{A^2\Pi-X^2\Sigma}$; ][]{Biver2018}.
%
However this is normally significantly weaker than the observed \ce{CN} emission, with only less than 20 comets.
%
Given the ubiquitous signature of \ce{CN} emission in Solar system comets, we have embarked on an archival search for this species in the \ac{feb} around \bp{}.
%
Detection of this species would provide a solid link between cometary bodies in our Solar system and the \bp{} system.

@@

For the study in this paper, we use all the publicly available data from the \ac{harps} and the ESO 3.6m telescope in Chile.

@@

We collated all publicly available HARPS observations of \bp{} obtained between October 2003 and March 2020.
%
Although these observations cover a large baseline, the time between visits varies between 1 day and $\sim$2.2 years.
%
Similarly, the number of observations per epoch spans a wide range, between 1 and 356 exposures, while the exposure times vary between 12 and 900 seconds.
%
This results in an average SNR across order 0 - 10 (order numbers for HARPS will follow the labelling in the fits header -- corresponding to 3780\,\AA and 4075\,\AA) per exposure between 1 and 430, and a nightly SNR between 63 and 2000.
%
To avoid potential edge effects from order merging,  we used the E2DS files downloaded directly from the ESO archive for our analysis.

\subsection{Initial data processing}
The analysis was started from the 2d extracted spectra (E2DS files).
%
Each order was treated separately, with the main focus on orders 3 and 7 as these orders contain the main \ce{CN} bandhead as well as the \ion{Ca}{ii} H line, which is used to identify the exocomets.
%
All the spectra were transformed into the stellar rest-frame and onto a common wavelength grid using linear interpolation. %For UVES, the \ce{Fe} \hl{NNNN angstrom} line was used to track global wavelength drifts between different exposures which were corrected for at the same time. To measure these, a Gaussian was fit to the line, and the velocity offset from the rest frame position determined.

@@

\subsection{Combining the spectra and flattening them}\label{sect:comb}

After the blaze-variation corrections, all the spectra within a night are combined in order to further boost the signal-to-noise and improve the detection limits.
%
Each individual spectrum was weighted using the average signal-to-noise ratio provided in the header for orders 0 and 11 (3780\,\AA{} - 4075\AA).%\hl{, and OO and PP} for HARPS and UVES respectively.

@@

After this post-processing the overall (residual) blaze profile and stellar spectrum still need to be removed.
%
Care must be taken to prevent the removal of (potential) narrow features from the circumstellar disk (as seen in e.g., the \ce{Fe} and \ion{Ca}{II} lines).
%
This is done by combining the nightly spectra without significant cometary activity (see Sect.~\ref{sect:FEBid}) and binning this stacked spectrum in wavelength in bins of 21 pixels.
%
Before binning, the spectrum in each bin was MAD (median absolute deviation) clipped to within $\pm$4.5 MAD of the median of the flux in the bin.
%
The clipped average of the fluxes within each bin was used.
%
Subsequently, the overall spectrum of the star and (residual) blaze was determined by cubic interpolation across the binned points, and divided out of the spectrum for each night.
%
As before, this was done separately pre- and post-fiber change.

@@

After the initial data analysis, we proceed to with our search for lines from \ce{CN}.
%
Although we are interested in lines from the exocomets, we first co-add the data from all the individual epochs, in the stellar rest-frame, to obtain a deep search for circumstellar \ce{CN} absorption.
%
These coadded spectra %for \ac{harps} and \ac{uves} 
are shown in Figure~\ref{fig:spec_CN_stellar_frame}.
%
No obvious feature from the \ce{CN} band jumps out, and we need to combine the signal from the lines within the \ce{CN} band to improve the detection limits.
%
Since the exocomets have a clear radial velocity with respect to \bp{}, and since each exocomet will have its own distinct velocity, we selected the 36 epochs with the strongest exocometary features, a seen in the \ion{Ca}{ii} H line.

\subsection{Identifying exocomets from \ion{Ca}{ii} H line}\label{sect:FEBid}
The comet features in the \ion{Ca}{ii} H-line ($\lambda=3968$ \AA) were identified after removing the overall stellar line-profile and avoiding the absorption from the circumstellar disk by excluding the central 8 km\,s$^{-1}$.
%
The deepest absorption feature in each observing night was selected, and the location and depth stored.
%
For the comet sample we selected features deeper than 60\% %\hl{and 40\%} for HARPS and UVES, respectively, 
while spectra with features below 15\% %for both HARPS and UVES 
were flagged as `low activity'.
%
The distribution of feature locations and depths is shown in Figure~\ref{fig:FEB_velocity}.

The spectra for the strong comet sample were coadded, and the combined spectrum for the HARPS data is shown in Figure~\ref{fig:spec_CN_comet_frame}.
%
As with the circumstellar \ce{CN}, there are no obvious lines from \ce{CN} visible, and we will need to combine multiple lines to improve the signal-to-noise ratio, and see if there is any \ce{CN} present.

\begin{figure}
    \begin{centering}
        \includegraphics[width=\columnwidth]{figures/spec_stack_CN_stellar.pdf}
        \caption{Plot of the stacked spectrum in the stellar restframe (blue) with a CN model for $T=$300K and N=10$^{14}$\,cm$^{-2}$ overplotted in orange. The large signal around 3860\,\AA is due to the residual from a circumstellar \ion{Fe}{I} line.}
        \label{fig:spec_CN_stellar_frame}
        \script{plot_spec_mean.py}
    \end{centering}
\end{figure}

\begin{figure}
    \begin{centering}
        \includegraphics[width=\columnwidth]{figures/spec_stack_CN_comet.pdf}
        \caption{Plot of the stacked spectrum in the stellar rest frame (blue) with a CN model for $T=$300K and N=10$^{14}$\,cm$^{-2}$ overplotted in orange. The large signal around 3860\,\AA is due to the residual from a circumstellar \ion{Fe}{I} line.}
        \label{fig:spec_CN_comet_frame}
        \script{plot_spec_mean.py}
    \end{centering}
\end{figure}

@@

%\subsection{Summary of \ce{CN} in comets}
%\ce{CN} is one of the brightest optically active species in cometary comae.
%
%It is a minor component of the gas coma, with a production rate by number $\sim 1/500$th that of water. 
As discussed above in the introduction, most studies of Solar system comets observe the \ce{CN} molecules via optical fluorescence. From measurements and simple physical arguments the \ce{CN} absorption/emission is optically thin everywhere.
This means a flux from part of the coma can immediately be converted to a column density using a photon scattering efficiency for the molecular band.

\ce{CN} in solar system comets comes from two sources.
Sublimation of nucleus near-surface ices releases \ce{HCN}, which then photo-dissociates into \ce{CN}.
This in turn eventually photo-dissociates into the component atoms.

@@

Since observations of Comet Halley \citep{AHearn1986}, it has also been known that dust grains ejected from the nucleus also directly release \ce{CN} into the coma.
However, \ce{CN} production in comets is usually dominated by nuclear ice sublimation \citep{Fray2005}. The resulting \ce{CN} coma will have a true space density and apparent column density distribution centrally peaked at the comet nucleus in this case.

For Solar system comets, measured \ce{CN} column densities are often converted to an overall production rate by the comet of $Q_{\ce{CN}}$ molecules/sec via an analytical model known as the ``Haser'' model \citep{Haser2020}
This model gives the on-sky column density as a function of projected radial distance form the nucleus using only 3 parameters; the effective destruction scale length $l_p$ of the parents of \ce{CN} (\ce{HCN}+dust), the photodissociation scale length $l_{\ce{CN}}$ of \ce{CN}, and the gas outflow velocity $v$. The Haser model is somewhat simplistic but it represents the data well for solar system comets observed at typical distances from the Sun (1--3 au).

@@

We take the observed activity of 1P / Halley in 1985 / 1986 as a reference point for our estimates.
%
Although most Solar system comets are observed to be are far less active than Halley, some Long-Period comets show activity in excess of Halley.
Given that we will be dominated by the most active comets at \bp{}, this seems reasonable as something that can be later scaled to the most active comets such as Hale-Bopp \citep{Schleicher2024}.
%
When Halley was at $\sim 1$ au from the Sun, it was measured to have $Q_{\ce{CN}}\simeq 10^{27}$ molecules/sec using Haser models \citep{AHearn1995}.
%
The assumed outflow velocity in the models was generally  $0.85 - 1$ \kms{} at this distance.

For solar system comets, the normal Haser scale lengths used are \citep{Cochran86,AHearn1995} :

$$lp=1.7\times10^4r_h^2\:{\rm km}$$
$$l_{\ce{CN}}=2-3\times10^5r_h^2\:{\rm km}$$

@@

The lifetime of \ce{HCN} for photodissociation into \ce{H}+\ce{CN} at 1~au has been calculated to be $(3.2-7.7)\times10^4$ sec, with lifetime of \ce{CN} dissociating into \ce{C}+\ce{N} being $(1.3-3.1)\times10^5$ sec \citep{Huebner92}, depending on solar activity levels.
%
These are consistent with the measured scale lengths above if the parents of \ce{CN} are ejected at $\sim 0.5$ km/sec, and we take into account the excess photodissociation energy giving \ce{CN} an extra kick of $\simeq1$ \kms{} \citep{Combi1980}.

@@

Photodissociation is  driven by the UV photon flux at the comet, which in turn is dominated by Ly-$\alpha$ line flux. From Ly-$\alpha$ measurements in \citet{Landsman93} the intrinsic Ly-$\alpha$ luminosity of A6V/A7V stars is a factor of $\sim10$ higher than for G2V stars.
%
Hence the photodissociation lifetime of \ce{CN}  at 1 au from \bp{} of an `exoHalley' will be approximately 10 times smaller i.e. $\sim 2\times10^4$ sec at 1 au.

In steady-state sublimation, the sublimation rate is proportional to the photon flux at the comet. For \bp{} $L=8.7L_\odot$, so assuming 
all \ce{CN} production scales with stellar luminosity, exoHalley would have a larger \ce{CN} production rate of $Q_{\ce{CN}}\sim 9\times10^{27}$ molecules/sec.
%
The total number of \ce{CN} molecules in the coma at 1~au at any given time would therefore be

$$N_{\ce{CN}}=Q_{\ce{CN}} \tau_{\ce{CN}}$$
$$N_{\ce{CN}}\sim 9\times10^{27} \cdot  2\times10^4 \sim 2\times10^{32}\:{\rm molecules}$$
